### 2015-09-17 Paper_Markdown.Rmd
Lots of new text in the introduction.  Moved many plots into their own files (in R/figures).  Added some methods text.
New references.

Cut libraries that were unused.

Added a 'null' model.  Just randomly samples from the data without respect to underlying data.

---
title: "Non-analogues in paleoecological reconstruction"
author: "Simon Goring *et al*."
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    pandoc_args: -V geometry:vmargin=1in -V geometry:hmargin=1in
  md_document:
    variant: markdown_github
  word_document:
    fig_height: 6
    fig_width: 6
csl: elsevier-harvard.csl
pandoc_args: -H margins.sty
bibliography: non-analogue_clim.bib
---
Non-analogues in paleoecological reconstruction - the effects of analogue distance
========================================================

<!--

```{r setup-knitr, echo = FALSE, message = FALSE}
# Simon: Note that these are global options that apply to all chunks below.

knitr::opts_chunk$set(
  comment = " ",
  error = FALSE,
  # Note: I have set EVAL to false just to make the bib work. 
  # Once the bib is fixed, you'll set this to TRUE or just delete this line.
  eval = TRUE,
  cache = TRUE,
  tidy = FALSE,
  fig.path = "figure/",
  cache.path = "cache/"
)

library("pander")
library("plyr")
panderOptions('table.style', 'rmarkdown')
panderOptions('table.split.table', 300)
```
-->

Simon Goring^1^

J Sakari Salonen^2^

Miska Luoto^2^

Jack Williams^1,3^

^1^ Department of Geography, University of Wisconsin-Madison, Madison, WI, USA

^2^ Department of Geosciences and Geography, University of Helsinki, Helsinki, Finland

^3^ Center for Climatic Research, University of Wisconsin-Madison, Madison, WI, USA

**Abstract**
*Fossil pollen provides a widespread proxy for past vegetation.  It is increasingly used in paleoclimatic reconstruction, but the limits of its utility are not well known.  Methodological advances, such as the development of tests for spatial autocorrelation, and newer methods for climate reconstruction using boosted regression trees may improve the abilities of these climate reconstruction techniques, but little is known about the accuracy of these models under conditions of non-analogue vegetation known to occur at the end of the late glacial.  In this paper we generate non-analogue pollen assemblages by excluding close neighbours from calibration datasets at increasingly larger squared chord distances.  Using three standard and two newer techniques we then test the ability of pollen-based climate models to accurately recostruct climate under conditions of no-analogue vegetation.*

```{r LoadData, echo=FALSE, warning=FALSE, message=FALSE}

source('R/data_setup.R')

```

Introduction
=========================

Pollen assemblages are widely used for paleoclimatic reconstruction during the Holocene, but this method is also applied to pollen assemblages from the last Glacial period [@bartlein2011pollen, @kaufman2012multi], and earlier [@tarasov2011progress; @brewer2008climate, @cheddadi1998climate]. A particular challenge for climate reconstruction using these much earlier samples is the presence of non-analogue vegetation. Non-analogue pollen assemblages pose a problem for many reconstruction techniques, these vegetation assemblages may be the result of sets of climate variables that no longer co-exist, ecological conditions resulting from differing rates of species migration [@huntley1991plants] or release from herbivores [@gill2009pleistocene], or as a result of changes in human land use [@goring2015preset].

Predicting climate from pollen in non-analogue space is likely to increase the uncertainty of models, possibly introduce systematic bias in predictions, and is likely to increase the variability of predictions across a time-series of predictions from a single site. However, high-quality terrestrial proxies for climate are important for understanding past climate change, and can act as a constraint on models of climate change in the past derived from GCMs in regions where little proxy data is available.  Thus, understanding the behaviour of pollen-based climate reconstructions in non-analogue space is an important first stem in understanding these reconstructions.

Although we cannot know the values for temperature or precipitation during this time, statistical non-analogues do exist in a modern context, and we can construct false non-analogues using a modified h-block sampling procedure [*sensu*, @telford2009evaluation] that examines analogue distance rather than geographical distance.  The result of no-analogue prediction is likely to vary with the method used for climate reconstruction, and possibly the climate variable (or distribution of the climate variable) of interest. This has been shown to occur with reconstructions from mixed depositional environments [@goring2010pollen], and more broadly, related families of methods are likely to behave similarly [KNUTTIPaper, NEWCANNONPAPER].

The Modern Analogue Technique is effectively a k-nearest neighbors method, generally using a squared-chord metric of distance.  MAT is used commonly in paleoecological reconstruction [@overpeck1985quantitative], but the k-nearest neighbor alogrithm is common to big data analysis. With MAT, the effects of moving toward no-analogue space will likely vary depending on the location of the sample in multi-dimentsional space.  We expect that a sample close to the average of all samples will show increasing predictive variability, but little effect on the mean prediction as the closest analogues become progressvely further from the focal sample, but the average of distant samples remains similar. In general, with increasing analogue distance we expect samples should revert toward the mean, while also showing higher variability.  Of all the methods, MAT's dependence on the local neighborhood likely suggests that we should expect a higher increase in variance with increasing analogue distance.

When WA is used for reconstruction in no-analogue space it is likely that species combinations found along the climatic gradient of interest are absent, meaning higher variability of prediction and greater likelihood that species optima or tolerances will be misinterpreted.  This is likely to be most significant for taxa with bimodal distributions,

Methods:
=========================

Sites from across North America were used to build a modern analogue dataset.  Samples were obtained from the North American Modern Pollen Database east of 100^o^W.  Pollen taxa were standardized as in Gill et al. [@gill2015noanalogues], using a set of 54 taxa, similar to those reported in Whitmore et al. [-@whitmore2005modern].  We chose to exclude Western North America since pollen taxa such as *Tsuga canadensis* and *Tsuga heterophylla* may be morphologically indistinct, but their parent plant taxa have clearly differentiated distributions, thus it may be possible to find false close-analogues using the entire North American dataset.  While this is not, strictly speaking, a major problem for this analysis, it will raise baseline error rates for models which then cause us to underestimate the change in error as analogue distance is increased.

```{r Fig1SiteLocations, fig.width=7, fig.height=6, fig.cap='Figure: North American site locations', echo=FALSE, message=FALSE, warning=FALSE}

source('R/figures/figure1_map.R')

map_plot

```

**Figure 1**. *Location of sample plots used for analysis.  Samples are derived from a modern pollen dataset.*

To analyse the effects of non-analogue pollen on model skill we chose five pollen-based reconstruction methods from the literature and apply them to the modern pollen data from eastern North America.  The models are MAT [@overpeck1985quantitative], WA [@ter1993weighted], WAPLS [@ter1989inferring], BRT and randomForest.  Each of these models has been used previously in the pollen literature, and has strengths and weaknesses.  MAT is commonly applied to reconstruct climate, for pollen, diatoms, chironomids and other taxa.  WA has a long history in the literature, and the recent development of monotone deshrinking has apparently improved performance further (Simpson, 2014?), but WA continues to suffer from a basic assumption of unimodal taxon distributions along climate gradients.

WAPLS is a modification of WA, whereby the residuals for each component are again weighted and regressed and the number of components are chosen by minimizing the root mean squared error (RMSE) of the model [@ter1993weighted].

BRT [@de2007boosted] and randomForest [@breiman2001random] are ensemble machine-learning approaches which have seen increasing use in fossil pollen literature during the past few years [@goring2010pollen;@salonen2014reconstructing;@salonen2012north;@zhao2012application]. Both machine learning methods are based on creating an ensemble of regression tree models with all pollen taxa used as predictors and an environmental variable of choice as the response. The random forest creates an ensemble of independent tree models (here 500 trees), while using bootstrap samples of the calibration data when building each tree, and randomly selected subsets of the predictor set when determining each tree split, thus creating diversity in the tree ensemble. In BRT, by comparison, the tree ensemble is built sequentially, with each added tree aiming to explain the residuals of the previously fitted ensemble. The final BRT model can thus be likened to an additive regression model in which the individual terms are regression trees.

For each pollen model and each sample site we calibrated a model to reconstruct T_{jul} using a pool of samples selected by removing all pollen samples closer than an analogue distance that increased from 0 to 1 at a step size of 0.01.  The calibration dataset for each model and each sample, at each exclusion step was then bootstrapped 30 times, sampling with replacement from the available calibration samples, to obtain a variance parameter, along with a mean prediction, bias estimate, root mean squared error of prediction and sample size estimate.  This full procedure reqired 29,254,650 individual calibrated models (although a number failed due to small sample sizes with large analogue exclusion distances).  To facilitate this large number of models, and the significant processing time, we ran the simulations on an Amazon AWS EC2 server as a parallel process using 30 cores.

Results
=========================

When we examine the dissimilarity scores we can see that the distribution between minimum, maximum and overall dissimilarities can be understood numerically and spatially.

Mean minimum dissimilarity for pollen samples in the modern pollen dataset ranges from `r min(min.diss)` to `r max(min.diss)`, with a 95% confidence interval from `r quantile(min.diss, 0.025)` to `r quantile(min.diss, 0.975)`.  The highest minimum analogue distance is for an unpublished site along the New York, Pennsylvannia border.  The site is notable for very high proportions of *Fraxinus pennsylvanica/americana* pollen, more than 5x higher than the next closest site.  The lowest minimum distance occurs across four sites, two from Thinn Lake, identified as potentially anomalous in Williams et al [-@williams2009rapid] and XXXX.  The geographic distribution of non-analogue sites is particularly interesting since many of the sites are in close proximity to other modern sites.  This indicates that While spatial autocorrelation is likely to play a role in the proximity of analogues (and thus their predictions) [@telford2009evaluation] it is not always the case that spatially proximate samples are compositionally proximate, and indeed, pollen assemblages in densely sampled regions may exhibit no-analogue characteristics.

```{r Fig2MinDissim, fig.width=7, fig.height=6, fig.cap='Figure: North American site locations', echo=FALSE, message=FALSE, warning=FALSE}

dissim.scores <- data.frame(dissim = c(min.diss,
                                       as.vector(dists)[sample(length(dists), 100000)]),
                            class   = c(rep('Minimum', length(min.diss)),
                                        rep('All', 100000)))

ggplot(dissim.scores) + 
  geom_density(aes(x = dissim, linetype=class), alpha=0.5, size = 1.3, from = 0) +
  theme_bw() +
  xlab('Squared-chord Dissimilarity') +
  ylab('Kernel Density') +
  theme(text = element_text(size=24, family='serif'),
        strip.text = element_text(size=14, family='serif'),
        axis.text = element_text(size = 12, family='serif'),
        legend.position = 'none') +
  annotate("segment", x=0.48, 
           xend = 0.48, y = 0, yend = 0.1, size=2, color='black', alpha = 0.7) +
  scale_y_sqrt(limits = c(0, 5.5), expand=c(0,0)) +
  scale_x_continuous(expand=c(0,0))

```

**Figure 2**.  *Distribution of squared-chord dissimilarity values.  The 95% confidence interval for the minumum spanning distance in the North American Modern Pollen Database is 0.10 - 0.48.  We choose 0.48 (marked on the x axis) as our cut-off for 'non-analogue' distances in the database.*

```{r Fig3SiteMDS, fig.width=7, fig.height=6, fig.cap='Figure: North American site MDS scores', echo=FALSE, message=FALSE, warning=FALSE}

source('R/figures/figure3.R')
run_figure3(pol.mds, min.diss, map, climate)

```

**Figure 3**. *Location of non-analogue points in the modern pollen data, presented within a non-metric multidimensional scaling ordination (Panel 1) and geographically.  Interestingly, although a number of geographically isolated points exist in the dataset, the majority of non-analogue points exist within the most dense region of sampling, indicating that non-analogue sites are not ecologically stratified.*

Climate Variables
=========================

We used the `r nrow(climate)` records from the North American Modern Pollen Database and associated climate data to provide reconstructions of key climatic variables. The process of calculating the lambda/lambda ratio from RDA (ref) produces interesting relationships, perhaps impacted by the breadth of the available climate and pollen data across North America.

```{r echo = FALSE, message=FALSE, warning=FALSE}

#  Sakari, can you take a look at this and let me know what you think?  Am I doing this right, and what do you make of these results?  Are we simply using too large a dataset?

pol <- mod_pol[,7:ncol(mod_pol)]
clim <- climate[,4:ncol(climate)]

clim.ratio <- rep(NA, ncol(clim))

for(i in 1:ncol(clim)){
  test.struct <- rda(pol ~ clim[,i], na.action = na.omit)
  clim.ratio[i] <- test.struct$CCA$tot.chi / test.struct$CA$tot.chi
}

names(clim.ratio) <- colnames(clim)
best.clim <- round(clim.ratio[order(clim.ratio, decreasing = TRUE)]*100, 0)

```

Variable | Percent Variability
-------- | -------------------
`r names(best.clim)[1]` | `r best.clim[1]`
`r names(best.clim)[2]` | `r best.clim[2]`
`r names(best.clim)[3]` | `r best.clim[3]`
`r names(best.clim)[4]` | `r best.clim[4]`
`r names(best.clim)[5]` | `r best.clim[5]`
`r names(best.clim)[6]` | `r best.clim[6]`
`r names(best.clim)[7]` | `r best.clim[7]`

Key variables based on the $\frac{\lambda_1}{\lambda_2}$ using pCCA appear to be climate variables relating to mean monthly percent summer sunshine (variables preceded by an 's').  These variables are likely to be variables that are combinations of temperature and preciptiation variables, however this value is not linear, nor is it continuous, meaning that .  While Telford & Birks (2009) have criticised sunshine reconstructions (Freshette et al., 2008) in the past (and perhaps with good reason), the strength of the sunshine signal in the climate data is worth noting, but the limitations of CCA may impact our ability to detect and accurately assess its importance.

Model outputs
=========================

Once the climate variables of interest were established we began by bootstrapping the models for these climate variables.  From the total pool of samples we generate a calibration dataset by sampling with replacement.

```{r Fig4SampleLoss, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3, fig.width=6}
load('data/mat.res.RData')
load('data/wa.res.RData')
load('data/wapls.res.RData')
load('data/brt.res.RData')
load('data/rfor.res.RData')

vals <- c(seq(0, 1, by=0.01))

point.loss <- 1 - mat.res$sample_size[,49] / nrow(climate)

na.samples <- data.frame(site = rep(climate[,1], ncol(mat.res$bias)),
                         lat = rep(climate$LATDD, ncol(mat.res$bias)),
                         long = rep(climate$LONDD, ncol(mat.res$bias)),
                         loss = point.loss,
                         bin = factor(round(ecdf(point.loss)(point.loss), 1)),
                         size = as.vector(mat.res$sample_size),
                         distance = rep(seq(0, 1, by=0.01), 
                                        each = nrow(mat.res$sample_size)))

na.samples <- na.samples[!is.na(na.samples$size),]

loss <- ggplot(data = na.samples, aes(x=distance, y=size)) + 
  geom_path(aes(group=site, color = bin), alpha = 0.4) +
  scale_color_brewer(palette = 'PiYG', type = 'div') +
  xlab('Analogue Distance') + ylab('Available Samples') +
  theme(text = element_text(size=24, family='serif'),
        axis.text = element_text(size=14, family='serif'),
        legend.position = 'none') +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand=c(0,0)) +
  geom_segment(aes(x=0.48, 
                   xend = 0.48, 
                   y = min(mat.res$sample_size, na.rm=TRUE), 
                   yend = nrow(climate)), 
               size=1, color='red', linetype=2)

loc.loss <- data.frame(lat =  climate$LATDD,
                       long = climate$LONDD,
                       loss = point.loss,
                       bin  = factor(round(ecdf(point.loss)(point.loss), 1)))

loss.map <-  ggplot(data = data.frame(map), aes(long, lat)) + 
  geom_polygon(aes(group=group), color='blue', alpha=0.2) +
  geom_point(data = loc.loss, 
             aes(x = long, y = lat, color = bin), alpha=0.4, size = 3) +
  theme_bw() +
  xlab('Longitude') + ylab('Latitude') +
  theme(axis.text=element_text(size=12, family='serif'),
        panel.background = element_rect(fill = 'lightblue'),
        legend.position = 'none') +
  coord_map(projection='albers', lat0=35, lat1=55, xlim=c(-100, -60), ylim=c(35, 80)) +
  scale_color_brewer(palette = 'PiYG', type = 'div')
  #scale_color_continuous(low = '#F08080', high='#4169E1')

grid.arrange(loss, loss.map, ncol=2)

```

**Figure 4**. *As the analogue distance exclusion is increased the available data set size is reduced, although some points continue to maintain a large number of potential analogues, this is directly related to the minimum distance to the closest analogue.  Points that maintain a large pool of analogues are far from neighbours.*

We have an interesting distribution of loss here.  First, the color coding is based on loss to 'non-analogue' conditions, and we see that it is possible for some samples to have relatively low loss to the no-analogue point, but then much higher losses to the extreme end of our analogue space (see the magenta lines at the very bottom end of Figure 4a.).  Secondly, the points that lose the most samples tend to be in the boreal forest.  To some degree this is surprising, however, the dominance of Pinus and Picea in these forests likely means that they have smaller analogue distances from one another than samples in the southern forests (with more diversity and therfore, likely fewer close analogues), but also, surpisingly, arctic samples also have higer inter-sample distances, and are thus less sensitive to losses of close analogues.

```{r, , echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=5}

get_analoguegradient <- function(dist){

  x <- findInterval(dist, vals)

  climate_stuff <- data.frame(tjul = c(climate$tjul, 
                                     brt.res$mean_prediction[,x],
                                     mat.res$mean_prediction[,x],
                                     rfor.res$mean_prediction[,x],
                                     wa.res$mean_prediction[,x],
                                     wapls.res$mean_prediction[,x]),
                            class = factor(c(rep('Actual', nrow(climate)),
                                             rep(c('BRT', 'MAT', 'rFor', 'WA', 'WAPLS'),
                                                 each = nrow(mat.res$mean_prediction)))),
                            dist = dist)

}

dist_dists <- na.omit(do.call(rbind.data.frame,lapply(c(0, .5, 1), get_analoguegradient)))

density_plot <- ggplot(dist_dists, aes(x = tjul, color = class, size = class, fill = class)) + 
  geom_density() +
  scale_size_manual(values = c(2, 1, 1, 1, 1, 1)) +
  scale_fill_manual(values = c('#d3d3d3', NA, NA, NA, NA, NA)) +
  theme_bw() +
  coord_cartesian(xlim=c(-4.2, 28.8)) +
  facet_wrap(~dist, ncol=1, scales='free_y')

density_plot

```
**Figure 5**. *Climate predictions for July temperature using the full data set at an analogue distance of 0, 0.5 and 1.  These results indicate that most predictive methods capture the full environmental gradient, however WA and WAPLS appear to over-emphasise the middle of the environmental gradient, BRT appears to truncate the gradient at both high and low ends.  As expected, MAT appears to capture the true gradient best when the full data set is used, but rapidly regresses to the mean as distance to the nearest neighbor increases.*

```{r Fig5PredictedValues, message=FALSE, echo=FALSE, warning=FALSE}

init.color <- colorRampPalette(c('#F08080', '#68d64f', '#4169E1'))
clim.colors <- seq(0, 31, length.out = 300)
init.set <- init.color(300)[findInterval(x = climate[,10], vec=clim.colors)]

model.out <- data.frame(site   = rep(rep(climate[,1], ncol(wa.res$bias)),5),
                        preds  = c(as.vector(mat.res$mean_prediction),
                                   as.vector(wa.res$mean_prediction),
                                   as.vector(wapls.res$mean_prediction),
                                   as.vector(rfor.res$mean_prediction),
                                   as.vector(brt.res$mean_prediction)),
                        bias   = c(as.vector(mat.res$bias),
                                   as.vector(wa.res$bias),
                                   as.vector(wapls.res$bias),
                                   as.vector(rfor.res$bias),
                                   as.vector(brt.res$bias)),
                        variance = c(as.vector(mat.res$variance),
                                   as.vector(wa.res$variance),
                                   as.vector(wapls.res$variance),
                                   as.vector(rfor.res$variance),
                                   as.vector(brt.res$variance)),
                        dissim = rep(rep(vals, each=nrow(wa.res$bias)),5),
                        model  = rep(c('MAT', 'WA', 'WAPLS', 'rFor', 'BRT'), 
                                     each = length(wa.res$bias)),
                        init = rep(rep(climate[,10], 101), 5))

#  This kills ~ 300,000 rows right now because of gaps in rfor.res and brt.res.
model.out <- model.out[!rowSums(is.na(model.out)) > 0,]

#  This is a temporary fix, 
sample.model <- function(x){
  #  How many samples exist for the model?
  set.len <- sum(model.out$model == x)
  if( set.len > 5000) set.len <- 5000

  set <- sample(which(model.out$model == x), set.len)
  start.end <- which(model.out$model == x & 
                       (model.out$dissim %in% c(0, 0.5, 1)))

  unique(c(set, start.end))

}

samp.set <- sapply(c('MAT', 'WA', 'WAPLS', 'rFor', 'BRT'), sample.model)

model.samp <- model.out[unlist(samp.set),]

#  This plot takes a really long time to plot.

ggplot(model.samp, aes(x = dissim, y = preds, group = site, color = init)) +
  geom_path(alpha = 0.1) +
  theme_bw() +
  facet_wrap(~model, nrow = 1) +
  scale_x_continuous(expand=c(0,0), breaks = c(0.5, 1)) +
  scale_y_continuous(expand=c(0,0), limits=c(0, 40)) +
  #scale_color_continuous(low = 'blue', high='pink', trans='sqrt') +
  theme(legend.position = 'none',
        axis.title = element_text(family = 'serif', face = 'bold', size = 14),
        strip.text = element_text(family = 'serif', face = 'bold', size = 14),
        axis.text = element_text(family = 'serif', size = 10),
        strip.text = element_text(family = 'serif', face = 'bold', size = 12)) +
  annotate("segment", x=0.48, 
           xend = 0.48, y = 0, yend = 100, size=1, linetype='dashed') +
  xlab('Calibration Dissimilarity') +
  ylab('Predicted Value')

```

**Figure 5**. *Change in site predictions over time.  Of great interest is the difference between model outputs as the exclusion zone increases.  MAT can never predict beyond the extent of the data samples in the pool, and so as the exclusion zone increases the predictions move toward the dataset mean.  WAPLS uses an initial transformation and then regresses against it to predict new values, so as the exclusion zone increases we see an increase in the spread of the predicted values, with much greater variability and a loss of accuracy.*

```{r Fig6BiasValues, message=FALSE, echo=FALSE, warning=FALSE}

ggplot(model.samp, aes(x = dissim, y = sqrt(bias), group = site, color = init)) +
  geom_path(alpha = 0.1) +
  theme_bw() +
  facet_wrap(~model, nrow = 1) +
  scale_x_continuous(expand=c(0,0), breaks = c(0.5, 1)) +
  scale_y_sqrt(expand=c(0,0)) +
  scale_color_continuous(low = 'blue', high='pink', trans='sqrt') +
  theme(legend.position = 'none',
        axis.title = element_text(family = 'serif', face = 'bold', size = 14),
        strip.text = element_text(family = 'serif', face = 'bold', size = 14),
        axis.text = element_text(family = 'serif', size = 10),
        strip.text = element_text(family = 'serif', face = 'bold', size = 12)) +
  xlab('Calibration Dissimilarity') +
  ylab('Sample Bias (RMSEP)')

```

**Figure 6**. *Change in model bias with increasing analogue distance.*

```{r Fig7VarianceValues, message=FALSE, echo=FALSE, warning=FALSE}

ggplot(model.samp, aes(x = dissim, y = sqrt(variance), group = site, color = init)) +
  geom_path(alpha = 0.1) +
  theme_bw() +
  facet_wrap(~model, nrow = 1) +
  scale_x_continuous(expand=c(0,0), breaks = c(0.5, 1)) +
  scale_y_sqrt(expand=c(0,0)) +
  scale_color_continuous(low = 'blue', high='pink', trans='sqrt') +
  theme(legend.position = 'none',
        axis.title = element_text(family = 'serif', face = 'bold', size = 14),
        strip.text = element_text(family = 'serif', face = 'bold', size = 14),
        axis.text = element_text(family = 'serif', size = 10),
        strip.text = element_text(family = 'serif', face = 'bold', size = 12)) +
  xlab('Calibration Dissimilarity') +
  ylab('Bootstrap Variance')

```

**Figure 7**. *Change in bootstrap variance with increasing analogue distance.*

Discussion
=========================
What does this tell us about predicting climate in non-analogue space?

Key points:  
Most predictive methods are stable until the no-analogue threshold is met.  Beyond this threshold the collapse of most methods is (surprisingly) rapid, in part due to the fact that the number of samples available for the calibration function drops substantially, indicating high sensitivity to the sample neighborhood for each of these methods.

* WAPLS can't be trusted given its bizarre behaviour beyond analogue space.

* MAT is incredibly dependent on finding close neighbors.  As analogue distance increases to near neighbours the likelihood of having poor fit increases substantially.  This is manifested in the apparent 'flip' in predictions.  We don't get regression to the mean, we get regression to the nearest available sample.
* BRT seems most robust, but for the purposes of this simulation (in lage part) it is mind-bogglingly slow, even when compared to rForest, howver (1) for a single run this is unlikely to be an issue (single runs can take somewhere between 1 and 10 minutes) and (2) interstingly BRT predictions are the most constrained, i.e., the predictive output has a much narrower range of predictions than the underlying data set.

References
=========================
Bartlein, P. J., Harrison, S. P., Brewer, S., Connor, S., Davis, B. A. S., Gajewski, K., ... & Wu, H. (2011). Pollen-based continental climate reconstructions at 6 and 21 ka: a global synthesis. Climate Dynamics, 37(3-4), 775-802.

Brewer, S., Guiot, J., Sánchez-Goñi, M. F., & Klotz, S. (2008). The climate in Europe during the Eemian: a multi-method approach using pollen data. Quaternary Science Reviews, 27(25), 2303-2315.